---
title: "Peak Bloom Prediction"
author: "Devendra Kumar Maindola"
date: "27/02/2022"
output:
  html_document:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      out.width = '80%')
```

## Instructions

In this Machine Learning Model(XG-Boost) we are predicting the peak bloom date in the coming decade for all four locations required by the competition.
We have combined data from 3 categories
1. Location Data - Latitude/Longitude of the site
2. Historic weather data - Temperature
3. Historic Peak Bloom date

Package Installation

```{r, eval=FALSE}

install.packages('tidyverse')

install.packages('xgboost')

install.packages('caret')

install.packages('lattice')

install.packages('DiagrammeR')


```

and then loaded via

```{r}
library(tidyverse)

library(xgboost)
library(caret)
library(lattice)
library(data.table)
library(DiagrammeR)

```


## Loading the data


The columns in the file(cherryblossomfeaturedata) are :

* Day_of_the_year - It indicated the day number. For Example Jan 1 = 1, feb 1 = 32
* Latitude 
* Longitude
* Month
* Year
* Max_Temperature - Max Temperature of the day in degrees Fahrenheit
* Avg_Temperature - Average Temperature of the day in degrees Fahrenheit
* Min_Temperature - Min Temperature of the day in degrees Fahrenheit
* GDD - Growing Degree Days. It is calculated as Avg_Temperature - 32. Negative value is considered as 0
* Peak_Bloom_Day - If Peak bloom date is 31st March, Peak Bloom Day will be 90
* Days to go for Peak Bloom - Number of days to the Peak Bloom event.

```{r}
df <- read.csv("data/cherryblossomfeaturedata.csv")

```
First 5 rows of the dataframe

```{r}

head(df,5)
```

## Keeping Year 2021 for Model Validation purpose 


df_valid <- filter(df, Year == 2021)

X_valid <- df_valid[ -c(6,7,8,10,11)]
Y_valid <- df_valid[ c(11)]

## Preparing Training and Test Dataset

```{r}
df_train <- filter(df, Year < 2021)

X_train <- df_train[-c(6,7,8,10,11)]

Y_train <- df_train[ c(11)]

df_test <- filter(df, Year > 2021)

X_test <- df_test[-c(6,7,8,10,11)]

Y_test <- df_test[ c(11)]


train_x <- data.matrix(X_train)
valid_x <- data.matrix(X_valid)
test_x <- data.matrix(X_test)

xgb_train <- xgb.DMatrix(data = train_x, label = Y_train[,1])
xgb_valid <- xgb.DMatrix(data = valid_x, label = Y_valid[,1])
xgb_test <- xgb.DMatrix(data = test_x, label = Y_test[,1])











## Predicting the peak bloom



# XG Boost Training
m_xgb_untuned <- xgboost(
    
    data = xgb_train,
    
    #n_estimators=1000,
    reg_lambda=1,
    gamma=1,
    max_depth=10,
    min_child_weight=4,
    learning_rate=0.05,
    
    nrounds = 1000,
    #reg_lambda=1,
    #gamma=1,
    objective = "reg:squarederror"
    
    #min_child_weight=4,
    #nfold = 5
    #max_depth = 20,
    #eta = 0.2
  )

# predict values in validation set for 2021


pred <- predict(m_xgb_untuned,valid_x)
pred[1]
pred[117]+58
pred[176]+59
pred[235]+59

MeanAbsoluteError = (85-pred[1] + 87-(58+pred[117]) + 87-(59+pred[176]) + 87-(59+pred[235]))/4
MeanAbsoluteError

```
##Observation : 1. 3.54 is the mean absolute error for all the sites(2021)
## 2. Jan1 is used for kyoto whreas feb end is used for other sites for the prediction 2021

## Prediction for 2022 onwards

pred1 <- predict(m_xgb_untuned,test_x)

d3 <- as.data.frame(pred1)
colnames(d3) <- c('days')

df_test$d3 <- d3
head(df_test,2)
df_35 <- filter(df_test,ï..Day_of_the_year==1 & Latitude==35.012)
df_35
df_other <- filter(df_test,ï..Day_of_the_year == 53 & Latitude!=35.012)


df_35$d3=round(df_35$d3,digits=0)

df_other$d3=round(df_other$d3,digits=0)
df_35 =select(df_35, Latitude,Longitude,Year,d3)

df_other =select(df_other, Latitude,Longitude,Year,d3)
df_35 <- transform(
  df_35, PeakBloomDayno= d3)

df_other <- transform(
  df_other, PeakBloomDayno= d3 + 53 )  
df_final <- rbind(df_35,df_other)
df_final <- df_final[-c(4)]

df_kyoto <- filter(df_final, Latitude == 35.0120)
df_liestal <- filter(df_final, Latitude == 47.4814)
df_washington <- filter(df_final, Latitude == 38.8853)
df_vancouver <- filter(df_final, Latitude == 49.2237)
df_kyoto <- df_kyoto %>%
  rename(kyoto=days)
df_liestal <- df_liestal %>%
  rename(liestal=days)
df_washington <- df_washington %>%
  rename(washingtondc=days)
  
df_vancouver <- df_vancouver %>%
  rename(vancouver=days)  
df_kyoto <- df_kyoto[-c(1,2)]
df_liestal <- df_liestal[-c(1,2)] 
df_washington <- df_washington[-c(1,2)] 
df_vancouver <- df_vancouver[-c(1,2)] 
submission <- cbind(df_kyoto,df_liestal,df_washington,df_vancouver)  
submission <- submission[-c(3,5,7)]
submission <- submission %>%
  rename(year=Year)


submission_predictions <- submission






write.csv(submission_predictions, file = "cherry-predictions.csv",
          row.names = FALSE)


xgb.plot.tree(model = m_xgb_untuned, trees = 1)




